Analysis, Prototyping and Locomotion Control of a Quadruped Robot\\

### 2023-06-17 LARS_paper/paper.tex
Fix images position and references order
Fix figure spacings, add title and ack
Fix .bib

\title{Conference Paper Title\\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
% should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.4\textwidth]{robot.jpeg}
  \vfill
  \caption{The quadruped robot.}
  \label{fig:robot}
  \vspace{-\baselineskip}
\end{figure}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{trot_pattern.png}
  \vfill
  % \vspace{-\baselineskip}
  \caption{Gait pattern diagram.}
  \label{fig:trot_pattern}
\end{figure}

@@

The trajectory planner is responsible for calculating the trajectory that each foot must perform. For both the stance and swing phases, a cycloidal curve was adopted. As shown in \cite{Shi2021}, a cycloidal curve can be defined in the three-dimensional space by (\ref{eq:traj_x}) to (\ref{eq:traj_k})
between the points $(x_o, y_o, z_o)$ and $(x_f, y_f, z_f)$ as a function of time $t$, step height $H$, and period $T$.
\begin{equation}
  x = (x_f - x_o) \frac{K - \sin{(K)}}{2 \pi} + x_o
  \label{eq:traj_x}
\end{equation}
\begin{equation}
  y = (y_f - y_o) \frac{K - \sin{(K)}}{2 \pi} + y_o
  \label{eq:traj_y}
\end{equation}
\begin{equation}
  z = H \frac{1 - \cos{(K)}}{2} + z_o
  \label{eq:traj_z}
\end{equation}
\begin{equation}
  K = \frac{2 \pi t}{T}
  \label{eq:traj_k}
\end{equation}

@@

\begin{figure*}[h]
  \centering
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_space.png}
    \caption{The cycloidal curve in 3D space.}
    \label{fig:traj_space}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_time.png}
    \caption{The cycloidal curve w.r.t. time.}
    \label{fig:traj_time}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_modified.png}
    \caption{The curve with an uneven distribution of points.}
    \label{fig:traj_time_modified}
  \end{subfigure}
  \vfill
  \caption{Cycloidal trajectories by the trajectory planner.}
  \label{fig:traj_curve}
\end{figure*}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4\textwidth]{caramel_tfs.png}

@@

\begin{figure}[htbp]
  \centering
  \vspace{-0.75cm}
  \includegraphics[width=0.35\textwidth]{caramel_body.drawio.png}

  \caption{Robot axes in resting position.}
  \label{fig:caramel_body}
\end{figure}

@@

The control subsystem of the robot is composed of two main components: the joints' controllers, and the body's stabilization controller. The joints' controllers are responsible for controlling each joint of the robot independently. Meanwhile, the body's stabilization controller is responsible for controlling the pitch and roll angles of the robot \cite{Shi2021, StanfordQuadruped}, maintaining it at 0 degrees.

%TODO refazer focando no controlador dos dynamixels
The joints' controller is the Joint Trajectory Controller, a controller of ROS which executes joint-space trajectories on a group of joints. According to \cite{JointTrajectory}, trajectories are specified as a set of waypoints to be reached at specific time instants, which the controller attempts to execute as well as the mechanism allows. In this work, the waypoints consist of positions and are reached using PID controllers loop-closed.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.48\textwidth]{PID.drawio.png}
  \caption{Designed PID controller.}
  \label{fig:pid}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Results of the gait trajectory tracking experiment.}
  \scalebox{1.3}{
  \begin{tabular}{ccc}
    \hline
    \textbf{Test} & \textbf{1}      & \textbf{2}  \\
    \hline
    $\bar{time} (s)$         & 0.557  & 0.557 \\
    \hline
    $\sigma_{time} (s)$       & 0.008  & 0.008 \\
    \hline
    $\bar{z}_{max} (cm)$       & 4.760  & 4.755 \\      
    \hline
    $\sigma_{z_{max}} (cm)$    & 0.023  & 0.028 \\
    \hline
    $\bar{x}_{final} (cm)$     & 5.034  & 3.024 \\
    \hline
    $\sigma_{x_{final}} (cm)$  & 0.009  & 0.013 \\
    \hline
    $\bar{y}_{final} (cm)$     & 2.924  & 4.915 \\      
    \hline
    $\sigma_{y_{final}} (cm)$  & 0.013  & 0.004 \\      
    \hline   
  \end{tabular}
  }
  \vspace{-\baselineskip}
  \label{tab:trajetoria}
\end{table}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{grafico_trajetoria_z.png}
  \caption{Trajectory executed by the foot.}
  \label{fig:grafico_trajetoria_xyz}
\end{figure}
\vspace{-\baselineskip}

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_plano.jpeg}
    \caption{Flat.}
    \label{fig:terreno_plano}
  \end{subfigure}
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_irregular.jpeg}
    \caption{Uneven.}
    \label{fig:terreno_uneven}
  \end{subfigure}
  \vfill
  \caption{Terrains used for the experiment.}
  \label{fig:terrenos}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Velocity and stability results.}
  \scalebox{1.3}{
  \begin{tabular}{ccccc}
    \hline
    \textbf{Test}              & \textbf{1}   & \textbf{2}  & \textbf{3}   & \textbf{4}   \\ \hline
    Ter.                       & Flat         & Flat        & Uneven       & Uneven       \\ \hline
    \begin{tabular}[c]{@{}c@{}}S. C. \end{tabular} & No           & Yes         & No           & Yes          \\ \hline
    \begin{tabular}[c]{@{}c@{}}Vel. \\ $(cm/s)$ \end{tabular} & 2.15         & 3.68        & 2.03         & 2.39         \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Vel}$  \\ $(cm/s)$ \end{tabular} & 0.040        & 0.075       & 0.016        & 0.047        \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Roll}$ \end{tabular} & 12.59\degree & 8.81\degree & 13.37\degree & 11.13\degree \\\hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Roll}$ \end{tabular} & 2.54\degree  & 2.36\degree & 0.78\degree  & 1.47\degree  \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Pitch}$ \end{tabular} & 11.55\degree & 8.31\degree & 15.19\degree & 12.09\degree \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Pitch}$ \end{tabular} & 1.89\degree  & 2.63\degree & 1.44\degree  & 1.66\degree  \\ \hline
  \end{tabular}
  }
  \label{tab:vel_stab}
\end{table}

In addition to the velocity results, the pitch and roll oscillations were compared to determine whether or not the stabilization controller lowered the robot's body oscillation. The Shapiro-Wilk test was applied to the data from all the tests, after removing outliers, and every test returned a p-value greater than 0.05, which indicates that the distributions followed a normal curve. The ANOVA was, then, applied to compare the roll and pitch oscillations from tests 1 and 2 (flat terrain). The same process was applied to the tests 3 and 4 as well. The ANOVA revealed that the roll and pitch distributions of the tests with and without stabilization control were significantly different. Table \ref{tab:vel_stab} shows that the mean value of the tests with stabilization control was less than the mean value of the tests without it, indicating that the stabilization controller lowered the robot's body oscillations in both terrain types. Fig. \ref{fig:imu_test} better pictures the data distribution. Despite the more scattered distribution presented, 3 out of 4 of the tests with stabilization control had its third quartile less than the first quartile of the distribution without the controller for the same terrain type, demonstrating that the controller lowered the oscillations in at least 75\% of the collected test samples.

\begin{figure}[!htb]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{plane_boxplot.png}
    \caption{Flat terrain.}
    \label{fig:imu_test_plane}
  \end{subfigure}
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{irregular_boxplot.png}
    \caption{Uneven terrain.}
    \label{fig:imu_test_irregular}
  \end{subfigure}
  \caption{Robot body's oscillation in both terrains.}
  \label{fig:imu_test}
\end{figure}

@@

% \section*{Acknowledgment}

% The preferred spelling of the word ``acknowledgment'' in America is without
% an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B.
% G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor
% acknowledgments in the unnumbered footnote on the first page.

@@

% For papers published in translation journals, please give the English
% citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
  \bibitem{X.134} X. Zhang, L. Lang, J. Wang, and H. Ma, ‘The quadruped robot locomotion based on force control’, Proceedings of the 2015 27th Chinese Control and Decision Conference, CCDC 2015, pp. 5440–5445, 2015.
  \bibitem{ROS2Humble} S. Macenski, T. Foote, B. Gerkey, C. Lalancette, and W. Woodall, ‘Robot Operating System 2: Design, architecture, and uses in the wild’, Science Robotics, vol. 7, no. 66, p. eabm6074, 2022.
  \bibitem{Shi2021} Y. Shi, S. Li, M. Guo, Y. Yang, D. Xia, and X. Luo, ‘Structural design, simulation and experiment of quadruped robot’, Applied Sciences (Switzerland), vol. 11, 2021.
  \bibitem{leotti2005comparaccao} V. B. Leotti, A. R. Birck, and J. Riboldi, ‘Comparação dos Testes de Aderência à Normalidade Kolmogorov-smirnov, Anderson-Darling, Cramer--Von Mises e Shapiro-Wilk por Simulação’, Anais do 11\textordmasculine Simpósio de Estatística Aplicada à Experimentação Agronômica, 2005.
  \bibitem{cano2012six} E. L. Cano, J. M. Moguerza, and A. Redchuk, Six sigma with R: statistical engineering for process improvement, vol. 36. Springer Science \& Business Media, 2012.
  \bibitem{CaramelRepo} B. S. de Alencar, F. M. S. M. Barreto, and L. L. Souza (2022), ‘Caramel’. Github.com. https://github.com/Brazilian-Institute-of-Robotics/bir-black-mouth. (Accessed Jun. 13, 2023).
  \bibitem{Kitano2016} S. Kitano, S. Hirose, A. Horigome, and G. Endo, ‘TITAN-XIII: sprawling-type quadruped robot with ability of fast and energy-efficient walking’, ROBOMECH Journal, vol. 3, pp. 1–16, 2016.
  \bibitem{Zhong2019} Y. Zhong, R. Wang, H. Feng, and Y. Chen, ‘Analysis and research of quadruped robot’s legs: A comprehensive review’, International Journal of Advanced Robotic Systems, vol. 16, 3 2019.
  \bibitem{de2006quadrupedal} P. G. De Santos, E. Garcia, and J. Estremera, ‘Quadrupedal locomotion: an introduction to the control of four-legged robots’, vol. 1. Springer, 2006.
  \bibitem{StanfordQuadruped} N. Kau, A. Z. Schultz, and M. Bowers, ‘Stanford Quadruped’. Github.com. https://github.com/stanfordroboticsclub/StanfordQuadruped. (Accessed Jun. 13, 2023).
  \bibitem{JointTrajectory} ros2\_control Development Team. ‘Joint Trajectory Controller’. Control.ros.org. https://control.ros.org/foxy/doc/ros2\_controllers/joint\_trajectory\_contro
  ller/doc/userdoc.html. (Accessed Jun. 13, 2023).

  % 14   STANFORD Quadruped. Dispon ́ıvel em:〈https://github.com/stanfordroboticsclub/StanfordQuadruped〉. Acesso: 24 de nov de 2022.15   DIY hobby servos quadruped robot.Dispon ́ıvel em:〈https://hackaday.io/project/171456-diy-hobby-servos-quadruped-robot〉.Acesso: 24 de nov de 2022.
  % 15   DIY hobby servos quadruped robot.Dispon ́ıvel em:〈https://hackaday.io/project/171456-diy-hobby-servos-quadruped-robot〉.Acesso: 24 de nov de 2022.
  % 17   NOTSPOT robot simulation. Dispon ́ıvelem:〈https://github.com/lnotspotl/notspot_sim_cpp〉. Acesso: 24 de nov de 2022.

  % \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
  % \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
  % \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
  % \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
  % \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
  % \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
  % \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
% \vspace{12pt}
% \color{red}
% IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

>>>>

\title{Analysis, Prototyping and Locomotion Control of a Quadruped Robot\\
% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
% should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

@@

\begin{figure}[!tb]
  \centering
  \includegraphics[width=0.4\textwidth]{robot.jpeg}
  \vfill
  \caption{The quadruped robot.}
  \label{fig:robot}
  \vspace{-\baselineskip}
\end{figure}

@@

\begin{figure}[b]
  \vspace{-\baselineskip}
  \centering
  \includegraphics[width=0.45\textwidth]{trot_pattern.png}
  \vfill
  \caption{Gait pattern diagram.}
  \label{fig:trot_pattern}
\end{figure}

@@

The trajectory planner is responsible for calculating the trajectory that each foot must perform. For both the stance and swing phases, a cycloidal curve was adopted. As shown in \cite{Shi2021}, a cycloidal curve can be defined in the three-dimensional space by (\ref{eq:traj_x}) to (\ref{eq:traj_k})
between the points $(x_o, y_o, z_o)$ and $(x_f, y_f, z_f)$ as a function of time $t$, step height $H$, and period $T$.

\begin{equation}
  x = (x_f - x_o) \frac{K - \sin{(K)}}{2 \pi} + x_o
  \label{eq:traj_x}
\end{equation}
\begin{equation}
  y = (y_f - y_o) \frac{K - \sin{(K)}}{2 \pi} + y_o
  \label{eq:traj_y}
\end{equation}
\begin{equation}
  z = H \frac{1 - \cos{(K)}}{2} + z_o
  \label{eq:traj_z}
\end{equation}
\begin{equation}
  K = \frac{2 \pi t}{T}
  \label{eq:traj_k}
\end{equation}

@@

\begin{figure*}[h]
  \centering
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_space.png}
    \caption{The cycloidal curve in 3D space.}
    \label{fig:traj_space}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_time.png}
    \caption{The cycloidal curve w.r.t. time.}
    \label{fig:traj_time}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{Cycloid_modified.png}
    \caption{The curve with an uneven distribution of points.}
    \label{fig:traj_time_modified}
  \end{subfigure}
  \vfill
  \caption{Cycloidal trajectories by the trajectory planner.}
  \label{fig:traj_curve}
  \vspace{-\baselineskip}
\end{figure*}

@@

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.4\textwidth]{caramel_tfs.png}

@@

\begin{figure}[tb]
  \centering
  \vspace{-0.75cm}
  \includegraphics[width=0.35\textwidth]{caramel_body.drawio.png}

  \caption{Robot axes in resting position.}
  \label{fig:caramel_body}
  \vspace{-1.3\baselineskip}
\end{figure}

@@

The control subsystem of the robot is composed of two main components: the joints' controllers, and the body's stabilization controller. The joints' controllers are the built-in Dynamixel position PID controllers, and they control each joint of the robot independently. Meanwhile, the body's stabilization controller is responsible for controlling the pitch and roll angles of the robot \cite{Shi2021, StanfordPupper}, maintaining it at 0 degrees.

@@

\begin{figure}[b]
  \vspace{-\baselineskip}
  \centering
  \includegraphics[width=0.48\textwidth]{PID.drawio.png}
  \caption{Designed PID controller.}
  \label{fig:pid}
\end{figure}

@@

\begin{table}[!b]
  \vspace{-\baselineskip}
  \centering
  \caption{Results of the gait trajectory tracking experiment.}
  \scalebox{1.3}{
  \begin{tabular}{ccc}
    \hline
    \textbf{Test} & \textbf{1}      & \textbf{2}  \\
    \hline
    $\bar{time} (s)$         & 0.557  & 0.557 \\
    \hline
    $\sigma_{time} (s)$       & 0.008  & 0.008 \\
    \hline
    $\bar{z}_{max} (cm)$       & 4.760  & 4.755 \\      
    \hline
    $\sigma_{z_{max}} (cm)$    & 0.023  & 0.028 \\
    \hline
    $\bar{x}_{final} (cm)$     & 5.034  & 3.024 \\
    \hline
    $\sigma_{x_{final}} (cm)$  & 0.009  & 0.013 \\
    \hline
    $\bar{y}_{final} (cm)$     & 2.924  & 4.915 \\      
    \hline
    $\sigma_{y_{final}} (cm)$  & 0.013  & 0.004 \\      
    \hline   
  \end{tabular}
  }
  \label{tab:trajetoria}
\end{table}

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.48\textwidth]{grafico_trajetoria_z.png}
  \caption{Trajectory executed by the foot.}
  \label{fig:grafico_trajetoria_xyz}
  \vspace{-\baselineskip}
\end{figure}

@@

\begin{figure}[b]
  \centering
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_plano.jpeg}
    \caption{Flat.}
    \label{fig:terreno_plano}
  \end{subfigure}
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_irregular.jpeg}
    \caption{Uneven.}
    \label{fig:terreno_uneven}
  \end{subfigure}
  \vfill
  \caption{Terrains used for the experiment.}
  \label{fig:terrenos}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Velocity and stability results.}
  \scalebox{1.3}{
  \begin{tabular}{ccccc}
    \hline
    \textbf{Test}              & \textbf{1}   & \textbf{2}  & \textbf{3}   & \textbf{4}   \\ \hline
    Ter.                       & Flat         & Flat        & Uneven       & Uneven       \\ \hline
    \begin{tabular}[c]{@{}c@{}}S. C. \end{tabular} & No           & Yes         & No           & Yes          \\ \hline
    \begin{tabular}[c]{@{}c@{}}Vel. \\ $(cm/s)$ \end{tabular} & 2.15         & 3.68        & 2.03         & 2.39         \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Vel}$  \\ $(cm/s)$ \end{tabular} & 0.040        & 0.075       & 0.016        & 0.047        \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Roll}$ \end{tabular} & 12.59\degree & 8.81\degree & 13.37\degree & 11.13\degree \\\hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Roll}$ \end{tabular} & 2.54\degree  & 2.36\degree & 0.78\degree  & 1.47\degree  \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Pitch}$ \end{tabular} & 11.55\degree & 8.31\degree & 15.19\degree & 12.09\degree \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Pitch}$ \end{tabular} & 1.89\degree  & 2.63\degree & 1.44\degree  & 1.66\degree  \\ \hline
  \end{tabular}
  }
  \label{tab:vel_stab}
  \vspace{-\baselineskip}
\end{table}

In addition to the velocity results, the pitch and roll oscillations were compared to determine whether or not the stabilization controller lowered the robot's body oscillation. The Shapiro-Wilk test was applied to the data from all the tests, after removing outliers, and every test returned a p-value greater than 0.05, which indicates that the distributions followed a normal curve. The ANOVA was, then, applied to compare the roll and pitch oscillations from tests 1 and 2 (flat terrain). The same process was applied to the tests 3 and 4 as well. The ANOVA revealed that the roll and pitch distributions of the tests with and without stabilization control were significantly different. Table \ref{tab:vel_stab} shows that the mean value of the tests with stabilization control was less than the mean value of the tests without it, indicating that the stabilization controller lowered the robot's body oscillations in both terrain types. Fig. \ref{fig:imu_test} better pictures the data distribution. Despite the more scattered distribution presented, 3 out of 4 of the tests with stabilization control had its third quartile less than the first quartile of the distribution without the controller for the same terrain type, demonstrating that the controller lowered the oscillations in at least 75\% of the collected test samples.
\vspace{0.1\baselineskip}

\begin{figure}[!tb]
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{plane_boxplot.png}
    \caption{Flat terrain.}
    \label{fig:imu_test_plane}
  \end{subfigure}
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{irregular_boxplot.png}
    \caption{Uneven terrain.}
    \label{fig:imu_test_irregular}
  \end{subfigure}
  \caption{Robot body's oscillation in both terrains.}
  \label{fig:imu_test}
\end{figure}

@@

\section*{Acknowledgment}

The authors' acknowledgments.

@@

% For papers published in translation journals, please give the English
% citation first, followed by the original foreign-language citation \cite{b6}.
%----------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{Bibliography}
%CRITICAL: do not change the above two lines!!!
%----------------------------------------------------------

% \begin{thebibliography}{00}
%   \bibitem{X.134} X. Zhang, L. Lang, J. Wang, and H. Ma, ‘The quadruped robot locomotion based on force control’, Proceedings of the 2015 27th Chinese Control and Decision Conference, CCDC 2015, pp. 5440–5445, 2015.
%   \bibitem{ROS2Humble} S. Macenski, T. Foote, B. Gerkey, C. Lalancette, and W. Woodall, ‘Robot Operating System 2: Design, architecture, and uses in the wild’, Science Robotics, vol. 7, no. 66, p. eabm6074, 2022.
%   \bibitem{Kitano2016} S. Kitano, S. Hirose, A. Horigome, and G. Endo, ‘TITAN-XIII: sprawling-type quadruped robot with ability of fast and energy-efficient walking’, ROBOMECH Journal, vol. 3, pp. 1–16, 2016.
%   \bibitem{Zhong2019} Y. Zhong, R. Wang, H. Feng, and Y. Chen, ‘Analysis and research of quadruped robot’s legs: A comprehensive review’, International Journal of Advanced Robotic Systems, vol. 16, 3 2019.
%   \bibitem{de2006quadrupedal} P. G. De Santos, E. Garcia, and J. Estremera, ‘Quadrupedal locomotion: an introduction to the control of four-legged robots’, vol. 1. Springer, 2006.
%   \bibitem{Shi2021} Y. Shi, S. Li, M. Guo, Y. Yang, D. Xia, and X. Luo, ‘Structural design, simulation and experiment of quadruped robot’, Applied Sciences (Switzerland), vol. 11, 2021.
%   \bibitem{StanfordQuadruped} N. Kau, A. Z. Schultz, and M. Bowers, ‘Stanford Quadruped’. Github.com. https://github.com/stanfordroboticsclub/StanfordQuadruped. (Accessed Jun. 13, 2023).
%   \bibitem{JointTrajectory} ros2\_control Development Team. ‘Joint Trajectory Controller’. Control.ros.org. https://control.ros.org/foxy/doc/ros2\_controllers/joint\_trajectory\_contro
%   ller/doc/userdoc.html. (Accessed Jun. 13, 2023).
%   \bibitem{cano2012six} E. L. Cano, J. M. Moguerza, and A. Redchuk, Six sigma with R: statistical engineering for process improvement, vol. 36. Springer Science \& Business Media, 2012.
%   \bibitem{leotti2005comparaccao} V. B. Leotti, A. R. Birck, and J. Riboldi, ‘Comparação dos Testes de Aderência à Normalidade Kolmogorov-smirnov, Anderson-Darling, Cramer--Von Mises e Shapiro-Wilk por Simulação’, Anais do 11\textordmasculine Simpósio de Estatística Aplicada à Experimentação Agronômica, 2005.
%   \bibitem{CaramelRepo} B. S. de Alencar, F. M. S. M. Barreto, and L. L. Souza (2022), ‘Caramel’. Github.com. https://github.com/Brazilian-Institute-of-Robotics/bir-black-mouth. (Accessed Jun. 13, 2023).

%   % 14   STANFORD Quadruped. Dispon ́ıvel em:〈https://github.com/stanfordroboticsclub/StanfordQuadruped〉. Acesso: 24 de nov de 2022.15   DIY hobby servos quadruped robot.Dispon ́ıvel em:〈https://hackaday.io/project/171456-diy-hobby-servos-quadruped-robot〉.Acesso: 24 de nov de 2022.
%   % 15   DIY hobby servos quadruped robot.Dispon ́ıvel em:〈https://hackaday.io/project/171456-diy-hobby-servos-quadruped-robot〉.Acesso: 24 de nov de 2022.
%   % 17   NOTSPOT robot simulation. Dispon ́ıvelem:〈https://github.com/lnotspotl/notspot_sim_cpp〉. Acesso: 24 de nov de 2022.

%   % \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
%   % \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
%   % \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
%   % \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
%   % \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
%   % \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
%   % \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
% \end{thebibliography}
% \vspace{12pt}
% \color{red}
% IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.
